Something about correlation functions

### 2021-06-29 paper-correlations.tex
Add a section about actual execution times for Directional module

The efficiency of \code{s2sep} is $O(n\:log\,n)$ and efficiency of
\code{s2insep} is $O(n^2)$ where $n$ is the length of the array.

@@

The edges are extracted by calculating $|S(array)|$ where $S$ is the Sobel
operator. \textbf{TODO: Cite our first paper}.

@@

Our algorithm for calculating $P(r)$ for a two-phase system is the following:
\begin{algorithmic}[1]
  \Procedure{poresize}{$array, nbins$}
    \State $d \gets dist(array)$ \Comment{$dist$ calculates
      $\mathcal{D}(\bm{x})$ for each element of $array$}
    \State $f(x) \gets x \ne 0$
    \State $positive \gets filter(d, f)$
    \State Bin values in $positive$ to $nbins$ bins to compute a histogram $H$
    \State \textbf{return} $H$
  \EndProcedure
\end{algorithmic}

\subsection{Chord length function}
The chord length function $p^{(i)}(\bm{r})d\bm{r}$ is defined as the probability of
finding a chord of length between $|\bm{r}|$ and $|\bm{r} + d\bm{r}|$ and
paralled to $\bm{r}$ in phase $i$. A chords is a line segment between
intersections of an infinitely long line with the interface of the phase
$i$. This function, again, can be calculated by cutting slices from a
multidimensional array. The algorithm is as follows:
\begin{algorithmic}[1]
  \Procedure{chordlength}{$array, phase, nbins$}
    \State $lengths \gets [\quad]$
    \State $ind \gets array = phase$ \Comment{$array$ is compared with $phase$
      element-wise}
    \State $edgephase \gets maximum(array) + 1$ \Comment{``bogus'' phase for the
      interface of the phase $phase$}
    \State $edge \gets edgephase \cdot [\text{Detect edges in $ind$}]$ \Comment{We
      use distance transform here}
    \State $combo = min(edge + array, edgephase)$ \Comment{Combine the original
      array and the interface for the phase $phase$}
    \ForAll{$direction \in \{\text{Requested directions}\}$}
      \ForAll{$slice \in \{\text{Slices parallel to $direction$}\}$}
        \State $len \gets 0$
        \State $startonedge \gets false$
        \ForAll{$x \in slice$}
          \If{$x = edgephase$}
            \Comment{Edge found}
            \If{$len > 1 \land startonedge$}
              \State $lengths \gets (len-1):lengths$
            \EndIf
            \State $len \gets 0$
            \State $startonedge \gets true$
          \ElsIf{$x \ne phase$}
            \State $startonedge \gets false$
          \EndIf
          \State $len \gets len + 1$
        \EndFor
      \EndFor
    \EndFor
    \State Bin elements in $lengths$ to $nbins$ bins to compute a histogram $H$
    \State \textbf{return} $H$
  \EndProcedure
\end{algorithmic}

@@

>>>>

The efficiency of \code{s2sep} is $O(n\log n)$ and efficiency of
\code{s2insep} is $O(n^2)$ where $n$ is the length of the array.

@@

The edges are extracted by calculating $|S(array)|$ where $S$ is the Sobel
operator. \textbf{TODO: Cite our first paper}. Algorithmic efficiency of these
algorithms are $O(n\log n)$ because it's implemented using \code{s2sep} and
filtering to extract edges is $O(n)$.

@@

Our algorithm for calculating $P(r)$ for a two-phase system is the following:
\begin{algorithmic}[1]
  \Procedure{poresize}{$array, nbins$}
    \State $d \gets dist(array)$ \Comment{$dist$ calculates
      $\mathcal{D}(\bm{x})$ for each element of $array$}
    \State $f(x) \gets x \ne 0$
    \State $positive \gets filter(d, f)$
    \State Bin values in $positive$ to $nbins$ bins to compute a histogram $H$
    \State \textbf{return} $H$
  \EndProcedure
\end{algorithmic}
Algorithmic efficiency is $O(n)$ where $n$ is the number of elements in the
input array.

\subsection{Chord length function}
The chord length function $p^{(i)}(\bm{r})d\bm{r}$ is defined as the probability of
finding a chord of length between $|\bm{r}|$ and $|\bm{r} + d\bm{r}|$ and
paralled to $\bm{r}$ in phase $i$. A chords is a line segment between
intersections of an infinitely long line with the interface of the phase
$i$. This function, again, can be calculated by cutting slices from a
multidimensional array. The algorithm which works in $O(n)$ time where $n$ is
the number of elements in the input is as follows:
\begin{algorithmic}[1]
  \Procedure{chordlength}{$array, phase, nbins$}
    \State $lengths \gets [\quad]$
    \State $ind \gets array = phase$ \Comment{$array$ is compared with $phase$
      element-wise}
    \State $edgephase \gets maximum(array) + 1$ \Comment{``bogus'' phase for the
      interface of the phase $phase$}
    \State $edge \gets edgephase \cdot [\text{Detect edges in $ind$}]$ \Comment{We
      use distance transform here}
    \State $combo = min(edge + array, edgephase)$ \Comment{Combine the original
      array and the interface for the phase $phase$}
    \ForAll{$direction \in \{\text{Requested directions}\}$}
      \ForAll{$slice \in \{\text{Slices parallel to $direction$}\}$}
        \State $len \gets 0$
        \State $startonedge \gets false$
        \ForAll{$x \in slice$}
          \If{$x = edgephase$}
            \Comment{Edge found}
            \If{$len > 1 \land startonedge$}
              \State $lengths \gets (len-1):lengths$
            \EndIf
            \State $len \gets 0$
            \State $startonedge \gets true$
          \ElsIf{$x \ne phase$}
            \State $startonedge \gets false$
          \EndIf
          \State $len \gets len + 1$
        \EndFor
      \EndFor
    \EndFor
    \State Bin elements in $lengths$ to $nbins$ bins to compute a histogram $H$
    \State \textbf{return} $H$
  \EndProcedure
\end{algorithmic}

@@

\section{Efficiency of our algorithms}

To measure efficiecy of our algorithms we use a machine with the following
configuration:
\begin{itemize}
\item \textbf{CPU}: AMD Ryzen 5 1600X running at 2200~MHz
\item \textbf{Memory}: 32~Gb DDR4-2400 memory
\item \textbf{GPU}: \textbf{Insert nvidia gpu here}
\end{itemize}

We generate two (resp. three) dimensional square (resp. cubic) binary arrays
with different number of elements. The arrays are filled with thresholded value
noise using \code{ValueNoise.jl} package and this code:
\begin{verbatim}
using ValueNoise

getdata(side, dims :: Integer) =
    getdata(side, Val(dims))

getdata(side, :: Val{2}) =
    [value_noise(5x/side,
                 5y/side,
                 0.0,
                 5, 34365) < 0.5
     for x in 1:side, y in 1:side]

getdata(side, :: Val{3}) =
    [value_noise(5x/side,
                 5y/side,
                 5z/side,
                 5, 34365) < 0.5
     for x in 1:side, y in 1:side, z in 1:side]
\end{verbatim}
An example of a two-dimensional image generated in this way is on
\cref{fig:valuenoise}. The function \code{getdata} generates the same image
sampled with different resolutions. For two-dimensional case we generate ten
arrays with the side of a square $s = 1000, 2000, \dots, 10000$ pixels and
measure execution time for each function. For three-dimensional case we measure
execution time for ten cubic arrays with the side $s = 50, 100, \dots, 500$. The
result for \code{Directional} module is on \cref{fig:timings}. All functions are
calculated with default values of optional arguments (that is: closed walls
boundary conditions and computation along axial directions).

\begin{figure}[ht]
  \centering
  \includegraphics[width=.3\textwidth]{images/timing-image.png}
  \caption[]{{\small An example of 2D image used in timing measurements}}
  \label{fig:valuenoise}
\end{figure}

\onecolumngrid
\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/time-2d.png}
    \caption[]{{\small Two-dimensional case}}
    \label{fig:timings-2d}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/time-3d.png}
    \caption[]{{\small Three-dimensional case}}
    \label{fig:timings-3d}
  \end{subfigure}
  \caption[]{\small Execution times for different functions from
    \code{Directional} module.}
  \label{fig:timings}
\end{figure*}
\twocolumngrid
